---
title: "diamonds"
author: "Devrim Nesipoðlu"
date: "4 Ocak 2018"
output: html_document
---

## Assignment 3: Diamonds Price Estimation
#          MEF - BDA 503


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)


```
#Diamonds dataset

This dataset contains the prices and other attributes for around 50.000 diamons. The features included are the following:

  * carat: Weight of the diamond

  * cut: Quality of the cut (Fair/Good/Very Good/ Premium/Ideal)

  * color: Diamond color, ranging from J (worst) to D (best)

  * clarity: A measurement of how clear the diamond is (I1 (worst), SI1, SI2, VS1, VS2, VVS1, VVS2, IF (best))

  * depth: Total depth percentage = z / mean(x, y) = 2 * z / (x + y) (43–79)

  * table: Width of top of diamond relative to widest point (43–95)

  * price: Price in US dollars ($326–$18,823)

  * x: Length in mm (0–10.74)
  * y: Width in mm (0–58.9)
  * z: Depth in mm (0–31.8)




```{r echo=TRUE}

library(rpart)
library(rpart.plot)
library(tidyverse)
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
library(ggvis)
library(dplyr)
library(caret)
```

<br><br>

#Exploratory Data Analysis


General information about Diamonds dataset:


```{r echo=TRUE}
data(diamonds)
str(diamonds)
```

```{r echo=TRUE}
dim(diamonds)
##result=>53940 observation  , 10 variable

summary(diamonds)
##result=>cut ,colar,clarity =>factors

```



```{r echo=TRUE}
df <- diamonds[,c(1,2,3,4,5,6,7)]
str(df)
```


```{r echo=TRUE}
head(df)

```


# Data Visualization

#ggplot2

ggplot2 is a plotting system for R, based on Leland Wilkinson's "The grammar of graphics". 

#qplot
The qplot function is a convenient wrapper for creating a number of different types of plots using a consistent scheme. It is recommended for simple scatterplots. This function creates a ggplot object which will receive the actual plot.

#geom
An important parameter is geom. It specifies the geometric objects that define the graph type. Geom values include:

  * point
  * smooth
  * boxplot
  * line
  * histogram
  * density
  * bar
  * jitter

using the pipe operator:

  diamonds %>% qplot(carat, price, data = .)

  To color these points, we just need to tell the function that the colors to be used depend on the cut feature, for example, by using the color argument.


# Create a scatterplot for comparing price based on diamond cut and color

```{r echo=TRUE}
theme_set(theme_bw())
ggplot(diamonds,aes(x=carat,y=price))+geom_jitter(aes(color=color,shape=cut))+
labs(title="comparison of carat with price based on diamond cut and color")
```

```{r echo=TRUE}
ggplot(diamonds, aes(x=carat, y=depth, color=clarity)) + geom_point(alpha=0.3) + geom_smooth() +theme_light()+
labs(title="Comparing diamond depth with carat and clarity")
```

```{r echo=TRUE}
ggplot(diamonds,aes(x=cut,fill=color))+geom_bar()
```

```{r echo=TRUE}
ggplot(diamonds,aes(x=price))+geom_density(aes(fill=factor(cut)),alpha=0.7)+labs(title="Price grouped by cut",x="Price",fill="Cut")
```

```{r echo=TRUE}
ggplot(diamonds,aes(x=clarity,y=depth))+geom_violin(aes(fill=cut))+theme_classic()+labs(title="How Cut impacts depth and clarity")
```

# Predicting diamonds Price with Regression model CART on entire dataset.

Regression Trees are part of the CART family of techniques for prediction of a numerical target feature. rpart package with its CART algorithms is used. 


```{r echo=TRUE}
m1 <- rpart(price ~ ., data=df, method ="anova")
m1

```


```{r echo=TRUE}

rpart.plot(m1, type=3, digits=3, fallen.leaves = TRUE)
```

# Build model on train set and predict on test

```{r echo=TRUE}
#
p1 <- predict(m1, df)

```


```{r echo=TRUE}
Mean_Absolute_Error <- function(actual, predicted) {mean(abs(actual - predicted))}

Mean_Absolute_Error

Mean_Absolute_Error(df$price, p1)
```

#References

R - Regression Trees - CART video of Jalayer Academy 
https://www.youtube.com/watch?v=MoBw5PiW56k&t=321s

For Data Visualization
https://www.kaggle.com/devisangeetha/data-visualization
