---
title: "TITANIC DISASTER "
author: "Devrim Nesipoğlu"
output: html_document
---

I will try to explore the Titanic disaster training set available from Kaggle.com. The data set consists of 1309 paasengers who rode aboard the Titanic.

1. Explore the variables

The first step in EDA is reading in data and then exploring the variables. 

Overview

The data has been split into two groups:

* training set (train.csv) for machine learning models.It will provide the outcome (also known as the “ground truth”) for each passenger and model will be based on “features” like passengers’ gender and class.

* test set (test.csv) will be used how well model performs on unseen data.It does not provide the
the ground truth for each passenger.

Gender_submission.csv, a set of predictions that assume all and only female passengers survive, as an example of what a submission file should look like.

DATA DICTIONARY

VariableDefinitionKey survival Survival 0 = No, 1 = Yes pclass Ticket class 1 = 1st, 2 = 2nd, 3 = 3rd sex Sex Age Age in years sibsp # of siblings / spouses aboard the Titanic parch # of parents / children aboard the Titanic ticket Ticket number fare Passenger fare cabin Cabin number embarked Port of Embarkation C = Cherbourg, Q = Queenstown, S = Southampton
Variable Notes
pclass: A proxy for socio-economic status (SES)
1st = Upper
2nd = Middle
3rd = Lower

age: Age is fractional if less than 1. If the age is estimated, is it in the form of xx.5

sibsp: The dataset defines family relations in this way...
Sibling = brother, sister, stepbrother, stepsister
Spouse = husband, wife (mistresses and fiancés were ignored)

parch: The dataset defines family relations in this way...
Parent = mother, father
Child = daughter, son, stepdaughter, stepson
Some children travelled only with a nanny, therefore parch=0 for them.

```{r}
##Load all the library required 

library('ggplot2')
library('caret') 
library('dplyr') 
library('randomForest') 
library('rpart')
library('rpart.plot')
library('car')
library('e1071')
library('ggthemes')
library('corrplot')
library('plyr')
```


```{r}
##Lets Load raw data in the orginal form by setting stringsAsFactors = F

train.tit <- read.csv('C:/Users/DEVRIM.YALCINKAYA/Documents/GitHub/pj-nesipoglud/files/Titanic_Data/train.csv', stringsAsFactors = F)
test.tit  <- read.csv('C:/Users/DEVRIM.YALCINKAYA/Documents/GitHub/pj-nesipoglud/files/Titanic_Data/test.csv', stringsAsFactors = F)
test.tit$Survived <- NA

##Combine both test and train
full_titanic <- rbind(train.tit, test.tit)
```

```{r}
##Check the structure
str(full_titanic)
```

Check the dimensions of the data

```{r}
dim(full_titanic)
```

The output shows us that data set of 1309 records and 12 columns. 

Several of the column variables are encoded as numeric data types (ints and floats) but a few of them are encoded as "object". 

Let's check the head of the data to get a better sense of what the variables look like:

Check the first 5 rows

```{r}
print(head(full_titanic,5))
```

We can see that this dataset is consists of numeric columns and columns with text data.

Let's look at a statistical summary of the variables with summary

After getting a sense of the data's structure, it is a good idea to look at a statistical summary of the variables with df.describe():


```{r}
summary(full_titanic)
```

```{R}
glimpse(full_titanic)

```



Missing value imputation

We can see from summary that Age, Fare, and Embarked have missing values, and that there is a large range in Fare. Naturally, Survived is missing for all test data rows.

```{r}
###is there any Missing obesrvation
colSums(is.na(full_titanic))
```


```{r}
####Empty data
colSums(full_titanic=='')
```


```{r}
names <- full_titanic$Name
title <-  gsub("^.*, (.*?)\\..*$", "\\1", names)

full_titanic$title <- title

table(title) 
```

```{r}
###MISS, Mrs, Master and Mr are taking more numbers

###Better to group Other titles into bigger basket by checking gender and survival rate to aviod any overfitting


full_titanic$title[full_titanic$title == 'Mlle']        <- 'Miss' 
full_titanic$title[full_titanic$title == 'Ms']          <- 'Miss'
full_titanic$title[full_titanic$title == 'Mme']         <- 'Mrs' 
full_titanic$title[full_titanic$title == 'Lady']          <- 'Miss'
full_titanic$title[full_titanic$title == 'Dona']          <- 'Miss'

## I am afraid creating a new varible with small data can causes a overfit
## However, My thinking is that combining below feauter into original variable may loss some predictive power as they are all army folks, doctor and nobel peoples 

full_titanic$title[full_titanic$title == 'Capt']        <- 'Officer' 
full_titanic$title[full_titanic$title == 'Col']        <- 'Officer' 
full_titanic$title[full_titanic$title == 'Major']   <- 'Officer'
full_titanic$title[full_titanic$title == 'Dr']   <- 'Officer'
full_titanic$title[full_titanic$title == 'Rev']   <- 'Officer'
full_titanic$title[full_titanic$title == 'Don']   <- 'Officer'
full_titanic$title[full_titanic$title == 'Sir']   <- 'Officer'
full_titanic$title[full_titanic$title == 'the Countess']   <- 'Officer'
full_titanic$title[full_titanic$title == 'Jonkheer']   <- 'Officer'


# Lets check who among Mr, Master, Miss having a better survival rate
 ggplot(full_titanic[1:891,],aes(x = title,fill=factor(Survived))) +
  geom_bar() +
  ggtitle("Title V/S Survival rate")+
  xlab("Title") +
  ylab("Total Count") +
  labs(fill = "Survived") 
```





```{r}
# Lets create a Family size using Sibsp and Parch

full_titanic$FamilySize <-full_titanic$SibSp + full_titanic$Parch + 1

full_titanic$FamilySized[full_titanic$FamilySize == 1]   <- 'Single'
full_titanic$FamilySized[full_titanic$FamilySize < 5 & full_titanic$FamilySize >= 2]   <- 'Small'
full_titanic$FamilySized[full_titanic$FamilySize >= 5]   <- 'Big'

full_titanic$FamilySized=as.factor(full_titanic$FamilySized)


###Lets Visualize the Survival rate by Family size 
ggplot(full_titanic[1:891,],aes(x = FamilySized,fill=factor(Survived))) +
  geom_bar() +
  ggtitle("Family Size V/S Survival Rate") +
  xlab("FamilySize") +
  ylab("Total Count") +
  labs(fill = "Survived")
```

Exploratory Analysis on Embarked

```{r}
###is there any association between Survial rate and where he get into the Ship.   
 ggplot(full_titanic[1:891,],aes(x = Embarked,fill=factor(Survived))) +
  geom_bar() +
  ggtitle("Embarked vs Survival") +
  xlab("Embarked") +
  ylab("Total Count") +
  labs(fill = "Survived") 
```


Age vs Survived


```{r}
ggplot(full_titanic, aes(Age, fill = factor(Survived))) + 
  geom_histogram(bins=30) + 
  theme_few() +
  xlab("Age") +
  scale_fill_discrete(name = "Survived") + 
  ggtitle("Age vs Survived")
```


Sex vs Survived

```{r}
# Sex vs Survived
ggplot(full_titanic, aes(Sex, fill = factor(Survived))) + 
  geom_bar(stat = "count", position = 'dodge')+
  theme_few() +
  xlab("Sex") +
  ylab("Count") +
  scale_fill_discrete(name = "Survived") + 
  ggtitle("Sex vs Survived")

tapply(full_titanic$Survived,full_titanic$Sex,mean)
```


Age vs Sex vs Survived


```{r}
#Sex vs Survived vs Age 
ggplot(full_titanic, aes(Age, fill = factor(Survived))) + 
  geom_histogram(bins=30) + 
  theme_few() +
  xlab("Age") +
  ylab("Count") +
  facet_grid(.~Sex)+
  scale_fill_discrete(name = "Survived") + 
  theme_few()+
  ggtitle("Age vs Sex vs Survived")
```


References

https://www.kaggle.com/vincentlugat/titanic-data-analysis-rf-prediction-0-81818/notebook
https://www.kaggle.com/swamysm/beginners-titanic

